---
title: "BA-ASSIGNMENT-01"
author: "ANURODH-SINGH"
date: "2023-10-03"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, results='hide'}
#Insertion/Addition of Package
options(repos = "https://cran.r-project.org/")
install.packages("ISLR")
library(ISLR)
```


```{r}
#Carseats is already an installed dataset in R library so seeing first of all what is Carseats data
print(Carseats)
```

```{r}
#Summary of dataset
summary(Carseats)
#Observations of dataset
nrow(Carseats)
```

```{r}
# maximum value of the advertising attribute
max(Carseats$Advertising)
```

```{r}
#Calculation of IQR
IQR(Carseats$Price)
```

```{r}
#Creation of Scatter Plot
x=Carseats$Sales
y=Carseats$Price
plot(x,y,main="Scatter Plot of Sales VS Price",xlab="sales",ylab="Price",col="red")

```


```{r}
#Co-Relation
cor(Carseats$Sales,Carseats$Price)

#The corelation gives the negative value so we can say that both Sales and Price trend to move in the opposite direction.
```


##The assignment is based on the creation and finding of the different analysis on the "Carseats" dataset.
#In this assignment we have uploaded the library called "ISLR".
#Next we have chosen the dataset called "Carseats".
#we have found out the summary for the dataset which is used to get the details.
# we have to find the maximum value using MAX .
#We found the Inter Quartile Range using IQR.
#Lastly,We have created a Scatterplot for Price VS Sales